\ldots

### 2016-02-08 figures/simlex-ppmi-best.tex
Complete the section on lexical tasks.

\begin{figure*}
  \centering
    \includegraphics[width=\textwidth]{supplement/figures/simlex-ppmi-best.pdf}

  \caption{\textbf{SPMI best configurations.}}
  \label{fig:simlex-ppmi-best}
\end{figure*}

>>>>

\begin{figure*}
  \centering
  \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{supplement/figures/SimLex999-ppmi-best.pdf}
    \caption{SimLex-999}
    \label{fig:simlex-ppmi-best-simlex}
  \end{subfigure}

  \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{supplement/figures/men-ppmi-best.pdf}
    \caption{MEN}
    \label{fig:simlex-ppmi-best-men}
  \end{subfigure}

  \caption{\textbf{SPMI best configurations.} Black line shows the best results reported by \protect\newcite{TACL570}.}
  \label{fig:simlex-ppmi-best}
\end{figure*}

### 2016-02-08 paper.tex
Complete the section on lexical tasks.

\newcite{TACL570} suggested to optimize PMI with a  set of parameters adopted from predictive models of \newcite{mikolov2013efficient}, most notably \emph{shifted PMI} abbreviated as \texttt{neg} and \emph{context distribution smoothing} referred to as \texttt{cds}. Their experiments and thus the parameter selection recommendations use highly dimensional vector spaces, namely word vectors have almost 200000 dimensions. In general, recent work on lexical distributional semantics with the results close to the state of the art involve vectors with a considerable dimensionality of hundreds of thousands \cite{baroni-dinu-kruszewski:2014:P14-1,kiela-clark:2014:CVSC}.

Contrary to the trend of employing highly dimensional vectors, older work on lexical distributional semantics and recent work on compositional distributional semantics employ vectors with much less dimensions, for example, 2000 \cite{Grefenstette:2011:ESC:2145432.2145580,kartsadrqpl2014,milajevs-EtAl:2014:EMNLP2014}, 3000 \cite{Dinu:2010:MDS:1870658.1870771,milajevs-purver:2014:CVSC} or 10000 \cite{polajnar-clark:2014:EACL,Baroni2010nouns}.

@@

PMI is not directly applicable to the co-occurrence counts because non-observed co-occurrences lead to infinite PMI values which in turn make it impossible to compute similarity. A straightforward ``fix'' to this problem is to replace all infinities with zeroes. Such a replacement preserve sparsity of the input co-occurrence matrix. Later in the paper, we use PMI to refer to a weighting with this fix.

@@

Context distribution smoothing (\texttt{cds}) is another parameter shown to influence performance of distributional models \cite{TACL570}. It smoothes the context distribution $P(x)$:
%
\begin{equation}
  \label{eq:cds}
  P_{\alpha}(x) = \frac{n(x)^{\alpha}}{\sum_{c}n(c)^{\alpha}}
\end{equation}
%
We experiment with values of 1 (absence of smoothing) and 0.75.

We also estimate probability of a context using the global context appearance in the corpus:
%
\begin{equation}
  \label{eq:cds-nan}
  P(x) = \frac{n(x)}{|C|}
\end{equation}
%
where $|C|$ is the size of the used corpus. We refer to this case as \textit{global} context probability.

@@

As context words we select 1000, 2000, 3000, 5000, 10000, 20000, 30000, 40000 and 50000 most frequent lemmatised nouns, verbs, adjectives and adverbs. All context words are part of speech tagged, but we do not make difference between refined word types, for example, we do not distinguish between transitive and ditransitive verbs.

@@

For tasks that involve composition, we test the following operators: addition
and multiplication \cite{mitchell2010composition}; Kronecker \cite{Grefenstette:2011:ETV:2140490.2140497}; Relational \cite{Grefenstette:2011:ESC:2145432.2145580}; Copy object and Copy subject \cite{Kartsaklis12aunified}; Frobenious addition, Frobenious multiplication and Frobenious outer \cite{kartsadrqpl2014}.

Table~\ref{tab:parameters} list the parameters and their values used in this work.

@@

\subsection{Shifted (positive) PMI performance}

@@

Average \SPMI/ performance increases for the dimensionality values in the range of 1K to 10K. After 10K dimensions, the average \SPMI/ performance degrades. \SPMI/ variance is the smallest for the 10K dimensional models and it grows as the dimension number increases afterwards. This suggests that there are particular parameter combinations that weaken or strengthen the results of the lowly and highly dimensional models.

@@

Figure~\ref{fig:simlex-ppmi-interaction} shows that 10K dimensional \SPMI/ is, indeed, the least sensitive to parameter selection. For models with dimensionality greater than 10K, context distribution smoothing should be used with $\alpha = 0.75$. For models with dimensionality lower than 10K, it is beneficial to use global context probabilities. Shifting also depends on the dimensionality, models with up to 3K dimensions should set $k = 0.5$, models with up to 20k dimensions should set $k = 1.4$ and models with a higher number of dimensions should set $k = 5$.

@@

>>>>

\newcite{TACL570} suggested to optimize PMI with a  set of parameters adopted from predictive models of \newcite{mikolov2013efficient}, most notably \emph{shifted PMI} abbreviated as \texttt{neg} and \emph{context distribution smoothing} referred to as \texttt{cds}. Their experiments and thus the parameter selection recommendations use highly dimensional vector spaces with word vector  dimensionality of almost \textbf{200K}. As such, recent work on lexical distributional semantics with the results close to the state of the art involve vectors with a considerable dimensionality of hundreds of thousands \cite{baroni-dinu-kruszewski:2014:P14-1,kiela-clark:2014:CVSC}.

Contrary to the trend of employing highly dimensional vectors, other work on lexical and compositional distributional semantics employ vectors with much less dimensions, for example, \textbf{2K} \cite{Grefenstette:2011:ESC:2145432.2145580,kartsadrqpl2014,milajevs-EtAl:2014:EMNLP2014}, \textbf{3K} \cite{Dinu:2010:MDS:1870658.1870771,milajevs-purver:2014:CVSC} or \textbf{10K} \cite{polajnar-clark:2014:EACL,Baroni2010nouns}.

@@

PMI is not directly applicable to the co-occurrence matrix because non-observed co-occurrences lead to infinite PMI values which in turn make it impossible to compute similarity. A straightforward ``fix'' to this problem is to replace all infinities with zeroes. Such a replacement preserve sparsity of the input co-occurrence matrix. Later in the paper, we use PMI to refer to a weighting with this fix.

@@

Context distribution smoothing (\texttt{cds}) is another parameter shown to influence performance of distributional models \cite{TACL570}. It smoothes the context distribution $P(x)$:
%
\begin{equation}
  \label{eq:cds}
  P_{\alpha}(x) = \frac{n(x)^{\alpha}}{\sum_{c}n(c)^{\alpha}}
\end{equation}
%
We experiment with values of 1 (absence of smoothing) and 0.75 and sometimes refer to such context probability estimation as \emph{local context probability}.

We also estimate probability of a context using the global context appearance in the corpus:
%
\begin{equation}
  \label{eq:cds-nan}
  P(x) = \frac{n(x)}{|C|}
\end{equation}
%
where $|C|$ is the size of the used corpus. We refer to this case as \emph{global context probability}.

@@

As context words we select 1K, 2K, 3K, 5K, 10K, 20K, 30K, 40K and 50K most frequent lemmatised nouns, verbs, adjectives and adverbs. All context words are part of speech tagged, but we do not make difference between refined word types, for example, we do not distinguish between transitive and ditransitive verbs.

@@

For tasks that involve composition, we test the following operators: \textbf{addition}
and \textbf{multiplication} \cite{mitchell2010composition}; \textbf{Kronecker} \cite{Grefenstette:2011:ETV:2140490.2140497}; \textbf{Relational} \cite{Grefenstette:2011:ESC:2145432.2145580}; \textbf{Copy object} and \textbf{Copy subject} \cite{Kartsaklis12aunified}; \textbf{Frobenious addition}, \textbf{Frobenious multiplication} and \textbf{Frobenious outer} \cite{kartsadrqpl2014}.

Table~\ref{tab:parameters} lists the parameters and their values used in this work.

@@

@@

Average \SPMI/ performance increases for the dimensionality values in the range of 1K to 10K. After 10K dimensions, the average \SPMI/ performance degrades. \SPMI/ score variance is the smallest for the 10K dimensional models and it grows as the dimension number increases afterwards. This suggests that there are particular parameter combinations that weaken or strengthen the results of the lowly and highly dimensional models.

@@

Figure~\ref{fig:simlex-ppmi-interaction} shows dimensionality-\texttt{cds} and dimensionality-\texttt{neg} parameter interactions. 10K dimensional \SPMI/ is, indeed, the least sensitive to parameter selection. For models with dimensionality greater than 10K, context distribution smoothing should be used with $\alpha = 0.75$. For models with dimensionality lower than 10K, it is beneficial to use global context probabilities. Shifting also depends on the dimensionality, models with up to 3K dimensions should set $k = 0.5$, models with up to 30K dimensions should set $k = 1.4$ and models with a higher number of dimensions should set $k = 2$.

@@

\NSPMI/ gives good results with local context probabilities ($\alpha = 1$), models with less than 20K dimensions should be used with $k = 1$, otherwise $k = 1.4$ should be preferred.

We use correlation similarity metric for \SPMI/ and \NSPMI/ based models, cosine similarity is used for \logNSPMI/.

The proposed heuristics work well for \logNSPMI/, where performance variance is low, the performance of models selected using the heuristics is not lower than 0.01 point of the best configuration. For \SPMI/ and \NSPMI/ this difference is much higher, refer to Figure~\ref{fig:simlex-ppmi-best-simlex} for more details.

Finally, to see whether the heuristics are applicable to other datasets, we contrast the selected models with the best scores on the MEN dataset (see Figure~\ref{fig:simlex-ppmi-best-men}). Again, \logNSPMI/ follows the best possible setup closely, especially for models with the dimension number greater than 10K. Surprisingly, the heuristics for \NSPMI/ are very effective on the MEN dataset, even though they are based on SimLex-999. Heuristics for \SPMI/ do not guarantee best possible results.

In general, model performance depends on vector dimensionality. It is beneficial to experiment with dense and unsmoothed low dimensional spaces ($k < 1$, global context probability); and sparse and smooth high dimensional spaces ($k > 1$, $\alpha = 0.75$). However, these heuristics do not guarantee the best possible result for \SPMI/ and \NSPMI/ because of high variance of the corresponding scores. Based on this we suggest to use \logNSPMI/ with the dimensionality of at least 20K to ensure good performance on lexical tasks.
